---
title: "food_waste_2019"
author: "Shelby Walsh"
date: "10/29/2020"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE, echo = FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(tidyr)
library(janitor)
library(kableExtra)
library(car)
library(devtools)
library(broom)
library(ggpubr)
library(ggbeeswarm)
library(RColorBrewer)
library(effsize)
library(stargazer)
```

Entering in 2019 Food waste data

```{r}
#only selecting items with a % discount code, these items are diverted foodwaste
item_discounts_2019 <- read_csv("item_discoun_food_waste_2019.csv") %>% 
  clean_names() %>%  
  select(item, quantity, reason, amount) %>% 
  group_by(item) %>% 
  filter(grepl('% off', reason))

```


```{r}
#sum the amount of each item and the cost do we can see how many items were diverted for report, might need to add calculation based on normal nada prices
discount_totals <- item_discounts_2019 %>% 
  summarise(quantity = sum(quantity), 
            cost = sum(amount))
```

(1) Egg Production
```{r}
eggs_category_1 <- discount_totals %>% 
 filter(grepl('Eggs|eggs', item))

#none in discounted section: Include code for future usage

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Eggs|eggs', discount_totals$item),]

# Product Category Code
eggs_category_1 <- eggs_category_1 %>%
  add_column(prod_cat = 1)
```

(2) Cheese manufacturing
```{r}
cheese_category_2 <- discount_totals %>% 
 filter(grepl('Cheese,|Cheeze|Cream Cheese|Feta|Shreds|Vegan Cheese|Charmisan|Cashew Ricotta|Pumpkin Cheese|Blue Heron', item))

#Again, no cheese in discounted items, but including code for future calculations

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Cheese,|Cheeze|Cream Cheese|Feta|Shreds|Vegan Cheese|Charmisan|Cashew Ricotta|Pumpkin Cheese|Blue Heron', discount_totals$item),]

# Product Category Code
cheese_category_2 <- cheese_category_2 %>%
  add_column(prod_cat = 2)
```

(3) Meat (except poultry)
```{r}
meats_category_3 <- discount_totals %>% 
 filter(grepl('Two Rivers Meat|Pepperoni|Bratwurst|Sausage|Chorizo|Jerky', item))

#Again another empty category, code for future use. Also the deli category isn't really used for deli products, but more miscellaneous 50% discount items

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Two Rivers Meat|Pepperoni|Bratwurst|Sausage|Chorizo|Jerky', discount_totals$item),]

# Product Category Code
meats_category_3 <- meats_category_3  %>%
  add_column(prod_cat = 3)
```

(13) Repurposed and Upcycled Materials
```{r}
scrap_category_13 <- discount_totals %>% 
 filter(grepl('Mill Factor', item))

#horray! the first item since dairy products! Code will be used to subract Mill factor out of discount_total and give it its product category number

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Mill Factor', discount_totals$item),]

# Product Category Code
scrap_category_13  <- scrap_category_13   %>%
  add_column(prod_cat = 13)

#woo it worked!!
```

(4) Fabric products
```{r}
fabrics_category_4 <- discount_totals %>% 
 filter(grepl('Fabric|Filter|Bags|Bag |Bag,|Rag|Wrap|cloth|Cloth|Pads|Jersey,|Coozie', item))

#another dud category. code input for constancy sake

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Fabric|Filter|Bags|Bag |Bag,|Rag|Wrap|cloth|Cloth|Pads|Jersey,|Coozie', discount_totals$item),]

# Product Category Code
fabrics_category_4 <- fabrics_category_4  %>%
  add_column(prod_cat = 4)
```

(5) Flours & flour mixes
```{r}
flours_category_5 <- discount_totals %>% 
 filter(grepl('Flour,|Flour |Pancake|Arrowroot|Corn Meal|Corn Starch|Tapioca Starch', item))

#again another goose egg category. My gut tells me that a majority of discounted products are milks, veggies, fruits, breads and condients. must continue forward with the code just incase

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Flour,|Flour |Pancake|Arrowroot|Corn Meal|Corn Starch|Tapioca Starch', discount_totals$item),]

# Product Category Code
flours_category_5 <- flours_category_5 %>%
  add_column(prod_cat = 5)

```

(6) Bread and bakery products
```{r}
#fingers crossed this actually picks up some of the 157 line items left
breads_category_6 <- discount_totals %>% 
 filter(grepl('Bread|Hamburger|Cheesebuns|Ciabatta|Crouton|Loaf|muffin', item))

#sweet beans! 8 more items accounted for! Now to subract them so there are les than 150 line items to go

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Bread|Hamburger|Cheesebuns|Ciabatta|Crouton|Loaf|muffin', discount_totals$item),]

# Product Category Code
breads_category_6 <- breads_category_6  %>%
  add_column(prod_cat = 6)

#highlight that scope_3 might have an error in breads
```

(7) Cookie, cracker, pasta, and tortilla
```{r}
cookie_category_7 <- discount_totals %>% 
 filter(grepl('Cookie|Macaron|Sinna|Nanaimo|Muffin|Brownie|Dough|Bakeshop|Cricket|Powerball|Crisp|Cracker|Tortilla|Pasta|Noodle|Rotini|Penne|Creste|Fusilli|Spaghetti, K|Spaghetti, D|Fettuccine|Lasagna|Ramen|Soba|Couscous', item))

#awesome 2 more down! lets get to subracting

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Cookie|Macaron|Sinna|Nanaimo|Muffin|Brownie|Dough|Bakeshop|Cricket|Powerball|Crisp|Cracker|Tortilla|Pasta|Noodle|Rotini|Penne|Creste|Fusilli|Spaghetti, K|Spaghetti, D|Fettuccine|Lasagna|Ramen|Soba|Couscous', discount_totals$item),]

# Product Category Code
cookie_category_7  <- cookie_category_7 %>%
  add_column(prod_cat = 7)
```

(8) Sugar and confectionery product manufacturing
```{r}
sugar_category_8 <- discount_totals %>% 
 filter(grepl('Molasses|Sugar,|Cacao|Cocoa|Chocolate,|Chocolate Chips|Hot Chocolate|Candycane|Truffles|Jelly Bean|Caramel|Delish Fish|Rings, Watermelon|Yogurt Raisins|Bark|Bears|Licorice|Worms', item))

#another empty category: insert extra code here

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Molasses|Sugar,|Cacao|Cocoa|Chocolate,|Chocolate Chips|Hot Chocolate|Candycane|Truffles|Jelly Bean|Caramel|Delish Fish|Rings, Watermelon|Yogurt Raisins|Bark|Bears|Licorice|Worms', discount_totals$item),]

# Product Category Code
sugar_category_8  <- sugar_category_8 %>%
  add_column(prod_cat = 8)
```

(9) Coffee and tea manufacturing
```{r}
coffee_category_9 <- discount_totals %>% 
 filter(grepl('Coffee|Tea,|Tea Co.|Amoda Tea|Tea Blend|Chai|Latte Mix|East Van Roasters', item))

#EMPTY! ugggggggggg

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Coffee|Tea,|Tea Co.|Amoda Tea|Tea Blend|Chai|Latte Mix|East Van Roasters', discount_totals$item),]

# Product Category Code
coffee_category_9   <- coffee_category_9 %>%
  add_column(prod_cat = 9)
```

(10) Fruit and vegetable canning, pickling, and drying
```{r}
dried_category_10 <- discount_totals %>% 
 filter(grepl('Dried|Raisins|Goji Berries,|Apple Rings|Figs|Pickle|Sauerkraut|Kimchi|Stemchi|Olives|Ginger Carrots|Mostarda|Bean Flakes|Veggie Broth Powder|Turkish Apricot', item))

#another 3 down! woot I can't wait for product category 26

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Dried|Raisins|Goji Berries,|Apple Rings|Figs|Pickle|Sauerkraut|Kimchi|Stemchi|Olives|Ginger Carrots|Mostarda|Bean Flakes|Veggie Broth Powder|Turkish Apricot', discount_totals$item),]

# Product Category Code
dried_category_10   <- dried_category_10 %>%
  add_column(prod_cat = 10)
```

(11)  Ice Cream and frozen dessert manufacturing
```{r}
icecream_category_11 <- discount_totals %>% 
 filter(grepl('Ice Cream|Earnest', item))

#zero the not so hero

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Ice Cream|Earnest', discount_totals$item),]

# Product Category Code
icecream_category_11   <- icecream_category_11 %>%
  add_column(prod_cat = 11)
```

(12) Frozen Food
```{r}
frozen_category_12 <- discount_totals %>% 
 filter(grepl('Frozen|Pierogi', item))

#another zero.... starting to regret picking this task lol

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Frozen|Pierogi', discount_totals$item),]

# Product Category Code
frozen_category_12   <- frozen_category_12 %>%
  add_column(prod_cat = 12)
```

(13) Flavoring Syrup and Concentrates
```{r}
syrup_category_13 <- discount_totals %>% 
 filter(grepl('Syrup|Concen', item))

#nothing

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Syrup|Concen', discount_totals$item),]

# Product Category Code
syrup_category_13   <- syrup_category_13 %>%
  add_column(prod_cat = 13)
```

(14) Dry, condensed, and evaporated dairy product manufacturing
```{r}
dairy_df_14 <-discount_totals %>% 
 filter(grepl('Milk Powder', item))

discount_totals <- discount_totals[!grepl('Milk Powder', discount_totals$item),]

# Product Category Code
dairy_df_14   <- dairy_df_14 %>%
  add_column(prod_cat = 14)
```

(15)Fluid Milk and butter manufacturing
```{r}
dairy_category_15 <- discount_totals %>% 
  filter(grepl('Milk|Mylk|Butter, Salted|Butter, Vegan|Cultured Butter|Cream|Yogurt|Smoothie', item))

#Sweet, 11 dairy products!! let's subract and keep pluggin'

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Milk|Mylk|Butter, Salted|Butter, Vegan|Cultured Butter|Cream|Yogurt|Smoothie', discount_totals$item),]

# Product Category Code
dairy_category_15   <- dairy_category_15 %>%
  add_column(prod_cat = 15)
```

(16)Soybean and other oilseed processing
```{r}
oilseed_category_16 <- discount_totals %>% 
 filter(grepl('Oil|Tofu|Tempeh|Tempea|Soy Protein|Pea Protein|Tahini|Sunflower Butter', item))

#we got 3 woot woot

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Oil|Tofu|Tempeh|Tempea|Soy Protein|Pea Protein|Tahini|Sunflower Butter', discount_totals$item),]

# Product Category Code
oilseed_category_16   <- oilseed_category_16 %>%
  add_column(prod_cat = 16)
```

(17) Oilseed farming
```{r}
oilseed_df_17 <- discount_totals %>% 
 filter(grepl('Lentil|Peas|Soybeans|Bean|, Raw|Cashews, Whole|Flax Seeds|Sunflower Seeds', item))

discount_totals <- discount_totals[!grepl('Lentil|Peas|Soybeans|Bean|, Raw|Cashews, Whole|Flax Seeds|Sunflower Seeds', discount_totals$item),]

# Product Category Code
oilseed_df_17   <- oilseed_df_17 %>%
  add_column(prod_cat = 17)
```

(18) Snack food manufacturing
```{r}
snack_df_18 <- discount_totals %>% 
  filter(grepl('Cashew|Almond|Peanut|Butter Cup|Pistachio|Sesame|Trail|Hummus|Homous|Dip|Tapenade|Chips|Sticks|Twists|Granola|Muesli|MUNCHeez', item))

discount_totals <- discount_totals[!grepl('Cashew|Almond|Peanut|Butter Cup|Pistachio|Sesame|Trail|Hummus|Homous|Dip|Tapenade|Chips|Sticks|Twists|Granola|Muesli|MUNCHeez', discount_totals$item),]

# Product Category Code
snack_df_18   <- snack_df_18 %>%
  add_column(prod_cat = 18)
```

(19) Grain farming
```{r}
grains_df_19 <- discount_totals %>% 
 filter(grepl('Rice|Wheat|Quinoa|Barley|Oats|Oat Bran|Amaranth|Buckwheat|Farro|Millet|Popcorn|Psyllium', item))

discount_totals <- discount_totals[!grepl('Rice|Wheat|Quinoa|Barley|Oats|Oat Bran|Amaranth|Buckwheat|Farro|Millet|Popcorn|Psyllium', discount_totals$item),]

# Product Category Code
grains_df_19   <- grains_df_19 %>%
  add_column(prod_cat = 19)

```

(20) Fishing
```{r}
fish_df_20 <- discount_totals %>% 
 filter(grepl('Skipper', item))

discount_totals <- discount_totals[!grepl('Skipper', discount_totals$item),]

# Product Category Code
fish_df_20   <- fish_df_20 %>%
  add_column(prod_cat = 20)
```

(21) Seasonings and dressings
```{r}
seasoning_df_21 <- discount_totals %>% 
 filter(grepl('Gathering Place|Sauce|naise|Vinegar|Mustard|Miso|Tamari|Bay Leaf|Beet Root|Garlic Powder|Onion Powder|Ginger Powder|Garam Masala|Salt|Pesto|Chili |Italian Seasoning|Granules|Paprika Powder|Vanilla Extract|Dressing|Ketchup|Burst', item))

discount_totals <- discount_totals[!grepl('Gathering Place|Sauce|naise|Vinegar|Mustard|Miso|Tamari|Bay Leaf|Beet Root|Garlic Powder|Onion Powder|Ginger Powder|Garam Masala|Salt|Pesto|Chili |Italian Seasoning|Granules|Paprika Powder|Vanilla Extract|Dressing|Ketchup|Burst', discount_totals$item),]

# Product Category Code
seasoning_df_21   <- seasoning_df_21 %>%
  add_column(prod_cat = 21)
```

(22) Breweries
```{r}
brewery_df_22 <- discount_totals %>% 
  filter(grepl('Kombucha|Dickie|Kefir', item))

discount_totals <- discount_totals[!grepl('Kombucha|Dickie|Kefir', discount_totals$item),]

# Product Category Code
brewery_df_22   <- brewery_df_22 %>%
  add_column(prod_cat = 22)
```

(23) Greenhouse, nursery, and floriculture production
```{r}
floriculture_df_23 <- discount_totals %>%
  filter(grepl('Flower|flowers', item))

discount_totals <- discount_totals[!grepl('Flower|flowers', discount_totals$item),]

# Product Category Code
floriculture_df_23   <- floriculture_df_23 %>%
  add_column(prod_cat = 23)
```

(24) Cleaning compounds
```{r}
cleaning_df_24 <- discount_totals %>%
  filter(grepl('Castile|Cleaner|Dish|Liquid|Toothpaste|Crush and Brush|Laundry Soda|Brightener|Borax|Bleach|Cleaning Powder', item))

discount_totals <- discount_totals[!grepl('Castile|Cleaner|Dish|Liquid|Toothpaste|Crush and Brush|Laundry Soda|Brightener|Borax|Bleach|Cleaning Powder', discount_totals$item),]

# Product Category Code
cleaning_df_24    <- cleaning_df_24  %>%
  add_column(prod_cat = 24)
```

(25) Toilet preparation manufacturing
```{r}
beauty_df_25  <- discount_totals %>%
  filter(grepl('Balm|Lotion|Gel|Sanitizer|Soap|Shampoo|Conditioner|Deodorant|Detox|Cleansing|3-in-1 Bar|Sunscreen|Epsom|Magnesium Chloride|Eucalyptus Steam|Cedar Tea Tree', item))

discount_totals <- discount_totals[!grepl('Balm|Lotion|Gel|Sanitizer|Soap|Shampoo|Conditioner|Deodorant|Detox|Cleansing|3-in-1 Bar|Sunscreen|Epsom|Magnesium Chloride|Eucalyptus Steam|Cedar Tea Tree', discount_totals$item),]

# Product Category Code
beauty_df_25 <- beauty_df_25  %>%
  add_column(prod_cat = 25)
```

(26) Fruit and tree nut farming
```{r}
fruit_df_26 <- discount_totals %>%
  filter(grepl('Apple|Apricot|Orange,|Oranges|Nectarine|Lemons|Grapefruit,|Tangelo|Peach|Plum|Pummelo|Persimmon|Banana|Avocado|Mango|Medjool|Coconut|Lime|Pear|berr|Nuts|Hazelnut|Walnut', item)) 

discount_totals <- discount_totals[!grepl('Apple|Apricot|Orange,|Oranges|Nectarine|Lemons|Grapefruit,|Tangelo|Peach|Plum|Pummelo|Persimmon|Banana|Avocado|Mango|Medjool|Coconut|Lime|Pear|berr|Nuts|Hazelnut|Walnut', discount_totals$item),]

# Product Category Code
fruit_df_26 <- fruit_df_26  %>%
  add_column(prod_cat = 26)
```


(27) Vegetable and melon farming
```{r}
vegetable_category_27 <- discount_totals %>% 
  filter(grepl('Broccoli|Bok|Brussel|Basil|Beet|Cabbage|Carrot|Corn|Celery|Cilantro|Chard|Cucumber|Dill|Kale|Eggplant|Garlic|Kohlrabi|Leek|Lettuce|Greens|Ginger|sqaush|Squash|Potato|Pepper|Tomato|Mushroom|Mint|Melon|melon|Cantaloupe|Honeydew|Onion|Parsley|Parsnip|Pumpkin|Rad|Rhubarb|Romaine|Rosemary|Rutabaga|Sage|Shallot|Spinach|Thyme|Turnip|Turmeric Root|Yams|Zucchini|Seeds', item))

#sweet non potatoes! we've made some impact! But the US and CEDA think potatoes are not veggies.... which I have some objection to!

# Subtract those words from discount_totals
discount_totals <- discount_totals[!grepl('Broccoli|Bok|Brussel|Basil|Beet|Cabbage|Carrot|Corn|Celery|Cilantro|Chard|Cucumber|Dill|Kale|Eggplant|Garlic|Kohlrabi|Leek|Lettuce|Greens|Ginger|sqaush|Squash|Potato|Pepper|Tomato|Mushroom|Mint|Melon|melon|Cantaloupe|Honeydew|Onion|Parsley|Parsnip|Pumpkin|Rad|Rhubarb|Romaine|Rosemary|Rutabaga|Sage|Shallot|Spinach|Thyme|Turnip|Turmeric Root|Yams|Zucchini|Seeds', discount_totals$item),]

# Product Category Code
vegetable_category_27   <- vegetable_category_27 %>%
  add_column(prod_cat = 27)
```

(28) Apparel manufacturing
```{r}
apparel_df_28 <- discount_totals %>%
  filter(grepl('Lunapads|Hair Ties', item))

discount_totals <- discount_totals[!grepl('Lunapads|Hair Ties', discount_totals$item),]

# Product Category Code
apparel_df_28   <- apparel_df_28 %>%
  add_column(prod_cat = 28)
```

(29) Glass Products
```{r}

#double check

glass_df_29 <- discount_totals %>%
  filter(grepl('Glass |Growler|Floss, Container|Deposit', item))

discount_totals <- discount_totals[!grepl('Glass |Growler|Floss, Container|Deposit', discount_totals$item),]

# Product Category Code
glass_df_29   <- glass_df_29 %>%
  add_column(prod_cat = 29)
```

(30) Cutlery and handtool manufacturing
```{r}
cutlery_df_30 <- discount_totals %>%
  filter(grepl('Cutlery|Spork|Straw', item))

discount_totals <- discount_totals[!grepl('Cutlery|Spork|Straw', discount_totals$item),]

# Product Category Code
cutlery_df_30   <- cutlery_df_30 %>%
  add_column(prod_cat = 30)
```

(31) Sanitary paper product manufacturing
```{r}
paper_df_31 <- discount_totals %>%
  filter(grepl('Toilet Paper', item))

discount_totals <- discount_totals[!grepl('Toilet Paper', discount_totals$item),]

# Product Category Code
paper_df_31   <- paper_df_31 %>%
  add_column(prod_cat = 31)
```

(32) Metal can, box, and other metal container (light gauge) manufacturing
```{r}
metal_df_32 <- discount_totals %>%
  filter(grepl('Onyx', item))

discount_totals <- discount_totals[!grepl('Onyx', discount_totals$item),]

# Product Category Code
metal_df_32   <- metal_df_32 %>%
  add_column(prod_cat = 32)
```

(33) Fiber, yarn, and thread mills
```{r}
thread_df_33 <- discount_totals %>%
  filter(grepl('Floss, Refill|Dryer Ball', item))

discount_totals <- discount_totals[!grepl('Floss, Refill|Dryer Ball', discount_totals$item),]

# Product Category Code
thread_df_33   <- thread_df_33 %>%
  add_column(prod_cat = 33)
```

(34) Animal production, except cattle and poultry and eggs
```{r}
honey_df_34 <- discount_totals %>%
  filter(grepl('Honey', item))

discount_totals <- discount_totals[!grepl('Honey', discount_totals$item),]

# Product Category Code
honey_df_34   <- honey_df_34 %>%
  add_column(prod_cat = 34)
```

(35) Other food items (except artificial sweeteners)
```{r}
otherfood_df_35 <- discount_totals %>%
  filter(grepl('Carob|Plant|Citric|Powder|Soda|Arame|Maple|Yeast|Salsa|TMRW|Wine|Seaweed', item))

discount_totals <- discount_totals[!grepl('Carob|Plant|Citric|Powder|Soda|Arame|Maple|Yeast|Salsa|TMRW|Wine|Seaweed', discount_totals$item),]

# Product Category Code
otherfood_df_35   <- otherfood_df_35 %>%
  add_column(prod_cat = 35)
```

(36) Other non-food items
```{r}
otherall_df_36 <- discount_totals %>%
  add_column(prod_cat = 36)
```

All categories brought back together
```{r}
all_categories <- rbind(
  eggs_category_1,
  cheese_category_2,
  meats_category_3,
  fabrics_category_4,
  flours_category_5,
  breads_category_6,
  cookie_category_7,
  sugar_category_8,
  coffee_category_9,
  dried_category_10,
  icecream_category_11,
  frozen_category_12,
  scrap_category_13,
  syrup_category_13,
  dairy_df_14,
  dairy_category_15,
  oilseed_category_16,
  oilseed_df_17,
  snack_df_18,
  grains_df_19,
  fish_df_20,
  seasoning_df_21,
  brewery_df_22,
  floriculture_df_23,
  cleaning_df_24,
  beauty_df_25,
  fruit_df_26,
  vegetable_category_27,
  apparel_df_28,
  glass_df_29,
  cutlery_df_30,
  paper_df_31,
  metal_df_32,
  thread_df_33,
  honey_df_34,
  otherfood_df_35,
  otherall_df_36
)
```

