---
title: "nada_scope_3"
author: "Shelby Walsh"
date: "10/27/2020"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE, echo = FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(tidyr)
library(janitor)
library(kableExtra)
#library(car) # this package is apparently my computer's sworn nemesis
library(devtools)
library(broom)
library(ggpubr)
library(ggbeeswarm)
library(RColorBrewer)
library(effsize)
library(stargazer)
```


# 1. Read in the data
```{r}
all_data <- read_csv("2019_vendor_invoice.csv") %>% 
  clean_names()
```

# 2. Select revelant columns and drop NAs for item.
```{r}
all_data <- all_data %>% 
  select(purchase_order, vendor, receipt_id, item, unit, date, cost_unit, total_quantity, total_cost) %>% 
  drop_na(item) 
```

# 3.Group suppliers by spend, and select the top suppliers:
```{r}
spend_grouping <- all_data  %>% 
  group_by(vendor) %>% 
  summarise(cost = sum(total_cost)) %>% 
  arrange(-cost) %>% 
  head(29)
# #29 is top 80%
# This is not relevant to what we decided to do.
```

# 4. Group products by spend.
```{r}
all_products <- all_data  %>% 
  group_by(item) %>% 
  summarise(quantity = sum(total_quantity), 
            cost = sum(total_cost))

# Remove non-product line items:
all_products <- all_products[!grepl('Deposit', all_products$item),]
```

# 5. The next step is sorting products into CEDA-related categories*

* When making changes, be sure to re-run all code chunks above to avoid sorting errors. (Ctrl+Alt+P)
** CEDA categories described here: https://docs.google.com/spreadsheets/d/1O_1TdRUn4D5utKODiZOS0G7OFIE8OhBaj53buOkoktc/edit#gid=0


(1) Egg Production
```{r}
# Eggs (Did not find any poultry in 2019 records)
eggs_category_1 <- all_products %>% 
 filter(grepl('Eggs|eggs', item))

# Subtract those words from all_products
all_products <- all_products[!grepl('Eggs|eggs', all_products$item),]

# Product Category Code
eggs_category_1 <- eggs_category_1 %>%
  add_column(prod_cat = 1)
```

(2) Cheese manufacturing
```{r}
# Cheese
cheese_category_2 <- all_products %>% 
 filter(grepl('Cheese,|Cheeze|Cream Cheese|Feta|Shreds|Vegan Cheese|Charmisan|Cashew Ricotta|Pumpkin Cheese|Blue Heron', item))

# Subtract those words from all_products
all_products <- all_products[!grepl('Cheese,|Cheeze|Cream Cheese|Feta|Shreds|Vegan Cheese|Charmisan|Cashew Ricotta|Pumpkin Cheese|Blue Heron', all_products$item),]

# Product Category Code
cheese_category_2 <- cheese_category_2 %>%
  add_column(prod_cat = 2)
```

(3) Meat (except poultry) Processing
```{r}
meats_category_3 <- all_products %>% 
 filter(grepl('Two Rivers Meat|Pepperoni|Bratwurst|Sausage|Chorizo|Jerky', item))
# Basically just what they get from Two Rivers Speciality Meats*
# * See note in CEDA database - Two Rivers is not just a processing facility, but a ranch as well

# Subtract those words from all_products
all_products <- all_products[!grepl('Two Rivers Meat|Pepperoni|Bratwurst|Sausage|Chorizo|Jerky', all_products$item),]

# Product Category Code
meats_category_3 <- meats_category_3  %>%
  add_column(prod_cat = 3)
```

(13) Repurposed and Upcycled Materials
```{r}
scrap_category_13 <- all_products %>% 
 filter(grepl('Mill Factor|Recycled Tote', item))

# Subtract those words from all_products
all_products <- all_products[!grepl('Mill Factor|Recycled Tote', all_products$item),]

# Product Category Code
scrap_category_13  <- scrap_category_13   %>%
  add_column(prod_cat = 13)
```

(4) Fabric products
```{r}
fabrics_category_4 <- all_products %>% 
 filter(grepl('Fabric|Filter|Bag |Bag,|Wrap|Dryer|cloth|Cloth|Pads|Jersey,', item))
# Note "bag" variations to exclude "baguette"

# Subtract those words from all_products
all_products <- all_products[!grepl('Fabric|Filter|Bag |Bag,|Wrap|Cloth|cloth|Pads|Jersey,', all_products$item),]

# Product Category Code
fabrics_category_4 <- fabrics_category_4  %>%
  add_column(prod_cat = 4)
```

(5) Flours & flour mixes
```{r}
# Flours and flour mixes
flours_category_5 <- all_products %>% 
 filter(grepl('Flour,|Flour |Pancake|Arrowroot|Corn Meal|Corn Starch|Tapioca Starch', item))
# Note "Flour" variations exclude "Flourgirl"

# Subtract those words from all_products
all_products <- all_products[!grepl('Flour,|Flour |Pancake|Arrowroot|Corn Meal|Corn Starch|Tapioca Starch', all_products$item),]

# Product Category Code
flours_category_5 <- flours_category_5 %>%
  add_column(prod_cat = 5)
```

(6) Bread and bakery products
```{r}
# Basically just what they get from Two Rivers Speciality Meats
breads_category_6 <- all_products %>% 
 filter(grepl('Bread|Croissant|Crouton|Loaf|Rye|Multigrain|Hamburger|buns|French Rolls|Yellow Basket|Pie Slice|Pie Filling', item))
# Hamburger retrieves "hamburger bun" without messing with anything sold by the 'bunch'
# Note that the next category includes cookies and crackers. This one is just breads.

# Subtract those words from all_products
all_products <- all_products[!grepl('Bread|Croissant|Crouton|Loaf|Rye|Multigrain|Hamburger|buns|French Rolls|Yellow Basket|Pie Slice|Pie Filling', all_products$item),]

# Product Category Code
breads_category_6 <- breads_category_6  %>%
  add_column(prod_cat = 6)
```

(7) Cookie, cracker, pasta, and tortilla
```{r}
cookie_category_7 <- all_products %>% 
 filter(grepl('Cookie|Macaron|Sinnabon|Nanaimo|Muffin|Dough|Bakeshop|Cricket|Powerball|Crisp|Cracker|Tortilla|Pasta|Noodle|Rotini|Penne|Creste|Fusilli|Spaghetti, K| Spaghetti, D|Fettuccine|Lasagna|Ramen|Soba|Couscous', item))
# "Macaron" works for macaron and macaroni
# Avoiding spaghetti squash

# Subtract those words from all_products
all_products <- all_products[!grepl('Cookie|Macaron|Sinnabon|Nanaimo|Muffin|Dough|Bakeshop|Cricket|Powerball|Crisp|Cracker|Tortilla|Pasta|Noodle|Rotini|Penne|Creste|Fusilli|Spaghetti, K| Spaghetti, D|Fettuccine|Lasagna|Ramen|Soba|Couscous', all_products$item),]

# Product Category Code
cookie_category_7  <- cookie_category_7 %>%
  add_column(prod_cat = 7)
```

(8) Sugar and confectionery product manufacturing
```{r}
sugar_category_8 <- all_products %>% 
 filter(grepl('Sugar,|Cacao|Cocoa|Chocolate,|Chocolate Chips|Hot Chocolate|Candycane|Truffles|Jelly Bean|Caramel|Delish Fish|Watermelon Rings|Yogurt Raisins|Bark|Bears|Licorice|Worms', item))
# Cocounut sugar popped up here too

# Subtract those words from all_products
all_products <- all_products[!grepl('Sugar,|Cacao|Cocoa|Chocolate,|Chocolate Chips|Hot Chocolate|Candycane|Truffles|Jelly Bean|Caramel|Delish Fish|Watermelon Rings|Yogurt Raisins|Bark|Bears|Licorice|Worms', all_products$item),]

# Product Category Code
sugar_category_8  <- sugar_category_8 %>%
  add_column(prod_cat = 8)
```

(9) Coffee and tea manufacturing
```{r}
coffee_category_9 <- all_products %>% 
 filter(grepl('Coffee|Tea,|Tea Co.|Tea Blend|Chai|Latte Mix|Matcha', item))

# Subtract those words from all_products
all_products <- all_products[!grepl('Coffee|Tea,|Chai|Latte Mix', all_products$item),]

# Product Category Code
coffee_category_9   <- coffee_category_9 %>%
  add_column(prod_cat = 9)
```

(10) Fruit and vegetable canning, pickling, and drying
```{r}
dried_category_10 <- all_products %>% 
 filter(grepl('Dried|Raisins|Goji Berries,|Apple Rings|Figs|Pickle|Sauerkraut|Kimchi|Stemchi|Olives|Ginger Carrots|Mostarda|Bean Flakes|Veggie Broth Powder|Turkish Apricot', item))

# Subtract those words from all_products
all_products <- all_products[!grepl('Dried|Raisins|Goji Berries,|Apple Rings|Figs|Pickle|Sauerkraut|Kimchi|Stemchi|Olives|Ginger Carrots|Mostarda|Bean Flakes|Veggie Broth Powder|Turkish Apricot', all_products$item),]

# Product Category Code
dried_category_10   <- dried_category_10 %>%
  add_column(prod_cat = 10)
```

(11) Ice cream and frozen dessert manufacturing
```{r}
icecream_category_11 <- all_products %>% 
 filter(grepl('Ice Cream|Earnest', item))

# Subtract those words from all_products
all_products <- all_products[!grepl('Ice Cream|Earnest', all_products$item),]

# Product Category Code
icecream_category_11   <- icecream_category_11 %>%
  add_column(prod_cat = 11)
```

(12) Frozen food
```{r}
frozen_category_12 <- all_products %>% 
 filter(grepl('Frozen|Pierogi', item))

# Subtract those words from all_products
all_products <- all_products[!grepl('Frozen|Pierogi', all_products$item),]

# Product Category Code
frozen_category_12   <- frozen_category_12 %>%
  add_column(prod_cat = 12)
```

(14) Dry, condensed, and evaporated dairy product manufacturing
```{r}
dairy_category_14 <- all_products %>% 
 filter(grepl('Milk Powder', item))

# Subtract those words from all_products
all_products <- all_products[!grepl('Milk Powder', all_products$item),]

# Product Category Code
dairy_category_14 <- dairy_category_14 %>%
  add_column(prod_cat = 14)
```

(15) Fluid milk and butter manufacturing
```{r}
dairy_category_15 <- all_products %>% 
 filter(grepl('Milk|Mylk|Butter, Salted|Butter, Vegan|Cultured Butter|Cream|Yogurt', item))

# Subtract those words from all_products
all_products <- all_products[!grepl('Milk|Mylk|Butter, Salted|Butter, Vegan|Cultured Butter|Cream|Yogurt', all_products$item),]

# Product Category Code
dairy_category_15 <- dairy_category_15 %>%
  add_column(prod_cat = 15)
```

(16) Soybean and other oilseed processing
```{r}
oilseed_category_16 <- all_products %>% 
 filter(grepl('Oil|Tofu|Tempeh|Tempea|Soy Protein|Pea Protein|Tahini', item))

# Subtract those words from all_products
all_products <- all_products[!grepl('Oil|Tofu|Tempeh|Tempea|Soy Protein|Pea Protein|Tahini', all_products$item),]

# Product Category Code
oilseed_category_16 <- oilseed_category_16 %>%
  add_column(prod_cat = 16)
```

(17) Oilseed farming
```{r}
oilseed_category_17 <- all_products %>% 
 filter(grepl('Lentil|Peas|Soybeans|Bean|, Raw|Cashews, Whole|Seeds,', item))

# Subtract those words from all_products
all_products <- all_products[!grepl('Lentil|Peas|Soybeans|Bean|, Raw|Cashews, Whole|Seeds,', all_products$item),]

# Product Category Code
oilseed_category_17 <- oilseed_category_17 %>%
  add_column(prod_cat = 17)
```

(18) Snack food manufacturing
```{r}
snack_category_18 <- all_products %>% 
  filter(grepl('Cashew|Almond|Peanut|Butter Cup|Pistachio|Sesame|Trail|Hummus|Homous|Dip|Tapenade|Chips|Sticks|Twists|Granola|Muesli', item)) %>% 
  add_column(prod_cat = 18)

# Subtract those words from all_products
all_products <- all_products[!grepl('Cashew|Almond|Peanut|Pistachio|Sesame|Trail|Hummus|Homous|Chips|Dip|Tapenade|Sticks|Twists|Granola|Muesli', all_products$item),]
```

(19) Grain farming
```{r}
grains_category_19 <- all_products %>% 
 filter(grepl('Rice|Wheat|Quinoa|Barley|Oats|Amaranth|Buckwheat|Farro|Millet|Popcorn|Psyllium', item)) %>% 
  add_column(prod_cat = 19)

# Subtract those words from all_products
all_products <- all_products[!grepl('Rice|Wheat|Quinoa|Barley|Oats|Amaranth|Buckwheat|Farro|Millet|Popcorn|Psyllium', all_products$item),]
```

(20) Fishing
```{r}
fish_category_20 <- all_products %>% 
 filter(grepl('Skipper', item)) %>% 
  add_column(prod_cat = 20)

# Subtract those words from all_products
all_products <- all_products[!grepl('Skipper', all_products$item),]
```

(21) Seasonings and dressings
```{r}
seasoning_category_21 <- all_products %>% 
 filter(grepl('Gathering Place|Sauce|naise|Vinegar|Mustard|Miso|Tamari|Bay Leaf|Beet Root|Garlic Powder|Onion Powder|Garam Masala|Salt|Pesto|Italian Seasoning|Granules|Paprika Powder', item)) %>% 
  add_column(prod_cat = 21)
# 'naise' captures vegenaise and mayonnaise

# Subtract those words from all_products
all_products <- all_products[!grepl('Gathering Place|Sauce|naise|Vinegar|Mustard|Miso|Tamari|Bay Leaf|Beet Root|Garlic Powder|Onion Powder|Garam Masala|Salt|Pesto|Italian Seasoning|Granules|Paprika Powder', all_products$item),]
```

(22) Breweries
```{r}
brewery_category_22 <- all_products %>% 
  filter(grepl('Kombucha|Dickie|Kefir', item)) %>% 
  add_column(prod_cat = 22)

# Subtract those words from all_products
all_products <- all_products[!grepl('Kombucha|Dickie|Kefir', all_products$item),]
```

(23) Greenhouse, nursery, and floriculture production
```{r}
floriculture_category_23 <- all_products %>%
  filter(grepl('Flower|flowers', item)) %>% 
  add_column(prod_cat = 23)

# Subtract those words from all_products
all_products <- all_products[!grepl('Flower|flowers', all_products$item),]
```

(24) Cleaning compounds
```{r}
cleaning_category_24 <- all_products %>%
  filter(grepl('Castile|Cleaner|Dish|Toothpaste|Crush and Brush|Laundry Soda|Brightener|Borax|Bleach|Laundry', item)) %>% 
  add_column(prod_cat = 24)

# Subtract those words from all_products
all_products <- all_products[!grepl('Castile|Cleaner|Dish|Toothpaste|Crush and Brush|Laundry Soda|Brightener|Borax|Bleach|Laundry', all_products$item),]
```

(25) Toilet preparation manufacturing
```{r}
beauty_category_25  <- all_products %>%
  filter(grepl('Balm|Lotion|Gel|Sanitizer|Soap|Shampoo|Conditioner|Deodorant|Detox|Cleansing|Sunscreen|Epsom|Magnesium Chloride|Eucalyptus Steam|Cedar Tea Tree', item)) %>% 
  add_column(prod_cat = 25)

# Subtract those words from all_products
all_products <- all_products[!grepl('Balm|Lotion|Gel|Sanitizer|Soap|Shampoo|Conditioner|Deodorant|Detox|Cleansing|Sunscreen|Epsom|Magnesium Chloride|Eucalyptus Steam|Cedar Tea Tree', all_products$item),]
```

(26) Fruit and tree nut farming
```{r}
fruit_category_23 <- all_products %>%
  filter(grepl('Apple|Apricots|Orange,|Oranges|Nectarine|Lemons|Grapefruit,|Tangelo|Peach|Plum|Pummelo|Banana|Avocado|Mango|Medjool|Coconut|Lime|Pear|berr|Nuts|Hazelnut|Walnut', item)) %>% 
  add_column(prod_cat = 26)

# Subtract those words from all_products
all_products <- all_products[!grepl('Apple|Apricots|Orange,|Oranges|Nectarine|Lemons|Grapefruit,|Tangelo|Peach|Plum|Pummelo|Banana|Avocado|Mango|Medjool|Coconut|Lime|Pear|berr|Nuts|Hazelnut|Walnut', all_products$item),]

```
(14) Tree nut Cocominos

(16) Oilseed farming
```{r}
test <- all_products %>% 
 filter(grepl('Hazelnut', item))
```

Tempea - Chickpea, Lentil
Peanut


Hemp 

