---
title: "transportation"
author: "Andrew Salvador"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidyr)
library(janitor)
library(kableExtra)
library(car)
library(devtools)
library(broom)
library(ggpubr)
library(ggbeeswarm)
library(RColorBrewer)
library(effsize)
library(stargazer)
```

# Read in the data 
```{r}
all_data <- read_csv("2019_vendor_invoice.csv") %>% 
  clean_names() 
```

Left Coast Naturals (1)
```{r}
distance1 <- all_data %>% 
  filter(grepl('Left Coast Naturals', vendor)) %>% 
  add_column(distance_km = 14)
  
```

Discovery Organics (2)
```{r}
distance2 <- all_data %>% 
  filter(grepl('Discovery Organics', vendor)) %>% 
  add_column(distance_km = 6.6)
```

Horizon Distributors (3)
```{r}
distance3 <- all_data %>% 
  filter(grepl('Horizon Distributors', vendor)) %>% 
  add_column(distance_km = 29.4)
```

Organic Matters (4)
```{r}
distance4 <- all_data %>% 
  filter(grepl('Organic Matters', vendor)) %>% 
  add_column(distance_km = 14)
```

LFT Group Brands Ltd. (5)
```{r}
distance5 <- all_data %>% 
  filter(grepl('LFT Group Brands Ltd.', vendor)) %>% 
  add_column(distance_km = 48.2)
```

Oneka Elements (6)
```{r}
distance6 <- all_data %>% 
  filter(grepl('Organic Matters', vendor)) %>% 
  add_column(distance_km = 1316)
```

Nelson Naturals	(7)
```{r}
distance7 <- all_data %>% 
  filter(grepl('Nelson Naturals', vendor)) %>% 
  add_column(distance_km = 1312)
```

Jiva Organics (8)
```{r}
distance8 <- all_data %>% 
  filter(grepl('Jiva Organics', vendor)) %>% 
  add_column(distance_km = 25.8)
```

Rehoboth Farm	(9)
```{r}
distance9 <- all_data %>% 
  filter(grepl('Rehoboth Farm', vendor)) %>% 
  add_column(distance_km = 178.6)
```

Dean's Milkman (10)
```{r}
distance10 <- all_data %>% 
  filter(grepl("Dean's Milkman", vendor)) %>% 
  add_column(distance_km = 105)
```

Westpoint Naturals (11)	
```{r}
distance11 <- all_data %>% 
  filter(grepl('Westpoint Naturals', vendor)) %>% 
  add_column(distance_km = 17.4)
```

Namasthé Tea Co. (12)
```{r}
distance12 <- all_data %>% 
  filter(grepl('Namasthé Tea Co.', vendor)) %>% 
  add_column(distance_km = 250)
```

Yoggu! (13)
```{r}
distance13 <- all_data %>% 
  filter(grepl('Yoggu!', vendor)) %>% 
  add_column(distance_km = 14.8)
```

Artisan Bake Shoppe	(14)
```{r}
distance14 <- all_data %>% 
  filter(grepl('Artisan Bake Shoppe', vendor)) %>% 
  add_column(distance_km = 21.4)
```


Onyx (15)
```{r}
distance15 <- all_data %>% 
  filter(grepl('Onyx', vendor)) %>% 
  add_column(distance_km = 15.2)
```

Hoochy ‘Booch Kombucha (16)
```{r}
distance16 <- all_data %>% 
  filter(grepl('Hoochy ‘Booch Kombucha', vendor)) %>% 
  add_column(distance_km = 4)
```

Brush Naked	(17)
```{r}
distance17 <- all_data %>% 
  filter(grepl('Brush Naked', vendor)) %>% 
  add_column(distance_km = 776)
```

East Van Roasters (18)
```{r}
distance18 <- all_data %>% 
  filter(grepl('East Van Roasters', vendor)) %>% 
  add_column(distance_km = 6.4)
```

Earnest Ice Cream	(19)
```{r}
distance19 <- all_data %>% 
  filter(grepl('Earnest Ice Cream', vendor)) %>% 
  add_column(distance_km = 6.4)
```

ULINE	(20)
```{r}
distance20 <- all_data %>% 
  filter(grepl('ULINE', vendor)) %>% 
  add_column(distance_km = 8310)
```

Plenty + Spare (21) YEEEEEEEEEEEEEET
```{r}
distance21 <- all_data %>% 
  filter(grepl("Plenty + Spare", vendor)) %>% 
  add_column(distance_km = 8.8)
```

Nellie's All-Natural (22)
```{r}
distance22 <- all_data %>% 
  filter(grepl("Nellie's All-Natural", vendor)) %>% 
  add_column(distance_km = 21.2)
```

The Modern Pantry	(23)
```{r}
distance23 <- all_data %>% 
  filter(grepl("The Modern Pantry", vendor)) %>% 
  add_column(distance_km = 21)
```

Dickie's Ginger	(24)
```{r}
distance24 <- all_data %>% 
  filter(grepl("Dickie's Ginger", vendor)) %>% 
  add_column(distance_km = 5.2)
```

Klippers Organics (25)
```{r}
distance25 <- all_data %>% 
  filter(grepl("Klippers Organics", vendor)) %>% 
  add_column(distance_km = 704)
```

A Bread Affair (26)
```{r}
distance26 <- all_data %>% 
  filter(grepl("A Bread Affair", vendor)) %>% 
  add_column(distance_km = 9.6)
```

Gathering Place Trading Company	(27)
```{r}
distance27 <- all_data %>% 
  filter(grepl("Gathering Place Trading Company", vendor)) %>% 
  add_column(distance_km = 528)
```

Fairware (28)
```{r}
distance28 <- all_data %>% 
  filter(grepl("Fairware", vendor)) %>% 
  add_column(distance_km = 4)
```

Nina's Pierogi (29)
```{r}
distance29 <- all_data %>% 
  filter(grepl("Nina's Pierogi", vendor)) %>% 
  add_column(distance_km = 83)
```


